Morphological diversity in tenrecs (Afrosoricida, Tenrecidae): Comparing tenrec skull diversity to their closest relatives

### 2015-04-13 paper/PeerJ_resubmission/Manuscript.tex
Figure and supplemental edits after PeerJ request
Update after PeerJ acceptance: edits to manuscript and new version of figure 2

\subsection{Data collection}
	One of us (SF) collected data from five museums: Natural History Museum, London (BMNH), Smithsonian Institute Natural History Museum, Washington D.C. (SI), American Museum of Natural History, New York (AMNH), Museum of Comparative Zoology, Cambridge M.A. (MCZ) and Field Museum of Natural History, Chicago (FMNH). We used the taxonomy in Wilson \& Reeder's Mammal Species of the World \citeyearpar{Wilson2005}, except for the recently discovered tenrec species \textit{Microgale jobihely} \citep{Goodman2006}. We photographed all of the intact tenrec and golden mole skulls available in the collections. This included 31 of the 34 species in the tenrec family \citep{Olson2013} and 12 of the 21 species of golden moles \citep{Wilson2005}.

@@

	Figure 2 depicts the landmarks and curves which we used for each skull view. For landmarks defined by dental structures, we used published dental sources where available \citep{Repenning1967, Eisenberg1969, Nowak1983, MacPhee1987, KnoxJones1992, Davis1997, Querouil2001, Nagorsen2002, Wilson2005, Goodman2006, Karatas2007, Hoffmann2008, Asher2008,  Muldoon2009, Lin2010}  to identify the number and type of teeth in each species. Detailed descriptions of the landmarks, as well as an example figure of landmarks on golden mole skulls, can be found in the supplementary material.

	When using semilandmark approaches there is a potential problem of over-sampling: simpler structures will require fewer semilandmarks to accurately represent their shape \citep{MacLeod2012}. To ensure that we applied a uniform standard of shape representation to each outline segment (i.e. that simple structures would not be over-represented and more complex features would not be under-represented), we followed the method outlined by MacLeod \citeyearpar{MacLeod2012} to determine the minimum number of semilandmark points which would give accurate representations of morphological shape. We used 54 points for skulls in dorsal view (10 landmarks, 44 semilandmarks across 4 curves), 73 points for skulls in ventral view (13 landmarks, 60 semilandmarks) and 44 points for skulls in lateral view (9 landmarks and 35 semilandmarks across 2 curves). See Figure 2 and the supplementary material for more details.

@@

	We tested the null hypothesis that tenrecs and golden moles have the same morphological diversity (the same mean Euclidean distance to the family centroid). If this is true, when we randomly assign the group identity of each species (i.e. shuffle the "tenrec" and "golden mole" labels) and then re-compare the morphological diversity of the two groups, there should be no significant difference between these results and those obtained when the species are assigned to the correct groupings.

	We performed this shuffling procedure (random assignation of group identity) 1000 times and calculated the difference in morphological diversity between the two groups for each permutation. This generated a distribution of 1000 values which are calculations of the differences in morphological diversity under the assumption that the null hypothesis (equal morphological diversity in the two families) is true. This method automatically accounts for differences in sample size because shuffling of the group labels preserves the sample size of each group: there will always be 12 species labelled as "golden mole" and then, depending on the analysis, either 31 or 17 species labelled as "tenrec". Therefore, the 1000 permuted values of differences in morphological diversity create a distribution of the expected difference in diversity between a group of sample size 31 (or 17 in the case of the tenrec data subset) compared to a group of sample size 12 under the null hypothesis that the two groups have the same morphological diversity. We compared the observed measures of the differences in morphological diversity between the two families to these null distributions to determine whether there were significant differences after taking sample size into account (two-tailed t test). Data and code for all of our analyses are available on GitHub \citep{Finlay2015c}.

@@

% Figure 2: Skulls: dorsal, ventral and lateral landmarks
% Landmarks (numbered points) and curves (outlines) for the skulls in dorsal,ventral and lateral view. See the supplementary material for detailed landmark descriptions. The skulls are an example of a \textit{Potamogale velox} (otter shrew tenrec), museum accession number BMNH 1934.6.16.2.

@@

% Figure 4: Morphospace (principal components) plot of morphological diversity in tenrec and golden mole skulls
% Principal components plots of the morphospaces occupied by tenrecs (triangles, n=31 species) and golden moles (circles, n=12 species) for skulls in dorsal (top left), ventral (top right) and lateral (bottom left) views. Each point represents the average skull shape of an individual species. Axes are principal component 1 (PC1) and principal component 2 (PC2) of the average scores from principal components analyses of mean Procrustes shape coordinates for each species.

>>>>

\subsection{Data collection}
	One of us (SF) collected data from five museums: Natural History Museum, London (BMNH), Smithsonian Institute Natural History Museum, Washington D.C. (SI), American Museum of Natural History, New York (AMNH), Museum of Comparative Zoology, Cambridge M.A. (MCZ) and Field Museum of Natural History, Chicago (FMNH). We used the taxonomy in Wilson \& Reeder's Mammal Species of the World \citeyearpar{Wilson2005}, except for the recently discovered tenrec species \textit{Microgale jobihely} \citep{Goodman2006}. We photographed all of the undamaged tenrec and golden mole skulls available in the collections. This included 31 of the 34 species in the tenrec family \citep{Olson2013} and 12 of the 21 species of golden moles \citep{Wilson2005}.

@@

	Figure 2 depicts the landmarks and curves which we used for each skull view. For landmarks defined by dental structures, we used published dental sources where available \citep{Repenning1967, Eisenberg1969, Nowak1983, MacPhee1987, KnoxJones1992, Davis1997, Querouil2001, Nagorsen2002, Wilson2005, Goodman2006, Karatas2007, Hoffmann2008, Asher2008,  Muldoon2009, Lin2010}  to identify the number and type of teeth in each species. Detailed descriptions of the landmarks can be found in the supplementary material (Tables S1, S2, S3) along with an example figure of landmarks on golden mole skulls (figure S1).

	When using semilandmark approaches there is a potential problem of over-sampling: simpler structures will require fewer semilandmarks to accurately represent their shape \citep{MacLeod2012}. To ensure that we applied a uniform standard of shape representation to each outline segment (i.e. that simple structures would not be over-represented and more complex features would not be under-represented), we followed the method outlined by MacLeod \citeyearpar{MacLeod2012}. This re-sampling method determines the minimum number of semilandmark points required to measure an outline length to at least 95\% accuracy of the true length of the outline. The procedure balances the need to represent outline shapes accurately without introducing error by over-sampling curves \citep{MacLeod2012}. We used 54 points for skulls in dorsal view (10 landmarks, 44 semilandmarks across 4 curves, Table S1), 73 points for skulls in ventral view (13 landmarks, 60 semilandmarks, Table S2) and 44 points for skulls in lateral view (9 landmarks and 35 semilandmarks across 2 curves, Table S3). See Figure 2 and the supplementary material for more details.

@@

	We tested the null hypothesis that tenrecs and golden moles have the same morphological diversity (the same mean Euclidean distance to the family centroid). If this is supported, when we randomly assign the group identity of each species (i.e. shuffle the "tenrec" and "golden mole" labels) and then re-compare the morphological diversity of the two groups, there should be no significant difference between these results and those obtained when the species are assigned to the correct groupings.

	We performed this shuffling procedure (random assignation of group identity) 1000 times and calculated the difference in morphological diversity between the two groups for each permutation. This generated a distribution of 1000 values which are calculations of the differences in morphological diversity under the assumption that the null hypothesis (equal morphological diversity in the two families) is true. This method automatically accounts for differences in sample size because shuffling of the group labels preserves the sample size of each group: there will always be 12 species labelled as "golden mole" and then, depending on the analysis, either 31 or 17 species labelled as "tenrec". Therefore, the 1000 permuted values of differences in morphological diversity create a distribution of the expected difference in diversity between a group of sample size N=31 (or N=17 in the case of the tenrec data subset) compared to a group of sample size N=12 under the null hypothesis that the two groups have the same morphological diversity. We compared the observed measures of the differences in morphological diversity between the two families to these null distributions to determine whether there were significant differences after taking sample size into account (two-tailed t test). Data and code for all of our analyses are available on GitHub \citep{Finlay2015c}.

@@

% Figure 2: Skulls: dorsal, ventral and lateral landmarks
% Landmarks (numbered points) and curves (outlines) for the skulls in dorsal (A),ventral (B) and lateral (C) view. See the supplementary material for detailed landmark descriptions. The skulls are an example of a \textit{Potamogale velox} (otter shrew tenrec), museum accession number BMNH 1934.6.16.2.

@@

% Figure 4: Morphospace (principal components) plot of morphological diversity in tenrec and golden mole skulls
% Principal components plots of the morphospaces occupied by tenrecs (triangles, n=31 species) and golden moles (circles, n=12 species) for skulls in dorsal (A), ventral (B) and lateral (C) views. Each point represents the average skull shape of an individual species. Axes are principal component 1 (PC1) and principal component 2 (PC2) of the average scores from principal components analyses of mean Procrustes shape coordinates for each species.
